Stochastic Wind Power Forecasting

### 2020-05-06 LaTeX_Code/Paper/paper_1line.tex
some reorganization

{\color{red} Change this paragraph. Discuss with the global dataset, the dataset after removing days when curtailing has been detected, the data set without curtailing and with Lamperti transform.}
We have available a year long data set from Uruguay based on 363 observation paths, each of which is 24-hours long with observations recorded every 10 min. In total, it is a data set of approximately fifty thousand data points recorded in 2019. See Figure (\ref{fig:sample_data}). Data are normalized with respect to the maximum power capacity of wind power production in Uruguay, which is $\SI{1474}{\mega\watt}$. However, sometimes the real production is artificially modifies due to curtailing. This effect becomes clear when we inspect for different power production levels. We split the available data into low, medium and high power range. See Figure (\ref{fig:data_curtailing}).\\
We identify and remove the days with curtailing, removing a total of 108 days. From the resulting 255 days, we use 127 days to train the system and 128 to test it. Also, to avoid correlation between days, we intercalate the days we use for training and testing ({\color{red} Show that 24h is enough to ensure independence}).

In this paper we present the phenomenological underlying model in Section \ref{Section_2} and describe the physical constraints in Section \ref{Section_3} and how these constrains can be met. Then, in Section \ref{Section_4}, we will introduce an alternative formulation of the model in Lamperti space. In Section \ref{Section_5}, we show our parameter estimation procedure and its results in Section \ref{Section_6}. We compare alternative models in Section \ref{Section_7} and different forecast providers in Section \ref{Section_8}.

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/5.eps}
\includegraphics[width=0.45\textwidth]{plots/245.eps}\\
\includegraphics[width=0.45\textwidth]{plots/661.eps}
\includegraphics[width=0.45\textwidth]{plots/805.eps}
\caption{Four samples from the Uruguayan data of 2019. Each sample compromises of two 24-hour paths. In blue is an hourly wind power production forecast. In orange is the actual wind power production recorded in 10 minute intervals.}
  \label{fig:sample_data}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/LP_6.eps}
\includegraphics[width=0.45\textwidth]{plots/MP_6.eps}\\
\includegraphics[width=0.45\textwidth]{plots/HP_6.eps}
\includegraphics[width=0.45\textwidth]{plots/AP_6.eps}
\caption{We see that forecast errors exhibit skewness, which is exaggerated as a consequence of curtailing in the production. Low power is when produced power is in $[0,0.3]$, mid-power is when it is in $(0.3,0.6]$, and high power when it is in $(0.6,1]$. We have a total of 363 days to use.}
  \label{fig:data_curtailing}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/LP.eps}
\includegraphics[width=0.45\textwidth]{plots/MP.eps}\\
\includegraphics[width=0.45\textwidth]{plots/HP.eps}
\includegraphics[width=0.45\textwidth]{plots/AP.eps}
\caption{We observe that skewness has been reduced after removing the days with curtailing. We have a total of 255 days with no curtailing.}
  \label{fig:data_after_clean}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/LP_t.eps}
\includegraphics[width=0.45\textwidth]{plots/MP_t.eps}\\
\includegraphics[width=0.45\textwidth]{plots/HP_t.eps}
\includegraphics[width=0.45\textwidth]{plots/AP_t.eps}
\caption{Transitions in error between forecast and real production. We can observe that they are not totally centered around zero.}
\label{fig:error_transitions}
\end{figure}

@@

%---BEGIN SECTION 2---
\section{Phenomenological  Model} \label{Section_2}

@@

%---BEGIN SECTION 3---
\section{State independent diffusion term: Lamperti transform} \label{Section_3}

Our model (\ref{VtSDE}) for the forecast error has a diffusion term that depends on the state variable $V_t$. To estimate the unknown model parameters, a recommended technique is to modify the SDE (\ref{VtSDE}) by applying the so-called Lamperti transform (see \cite[40--41]{iacus1}, \cite{moma}, \cite[98--100]{saso}) to the process $V$ to obtain a SDE for the transformed process whose diffusion term does not depend anymore on the state variable. To this purpose, we consider the following Lamperti transformation  
%{\color{red} (Remind that continuous-time models with different diffusion term are singular and the likelihood cannot be written down. Besides, it can be unstable to work with low frequency. If the frequency of the data is too low, the Lamperti transform helps to detect this problem (in principle, it is not clear how to determine the right frequency for estimation. )}
%\begin{equation}
%Z_t = h(V_t, t) = \frac{1}{\sqrt{2 \theta_0 \alpha}} \int_{\xi}^{x} \frac{1}{\sqrt{(u + p_t)(1 - u - p_t)}} \,du \Bigg{\vert}_{x = V_t} \,,
%\end{equation}
%where $\xi$ is an arbitrary point of the state space of the process $V$. The choice of $\xi = \frac{1}{2} - p_t$, where $p_t$ is a known input, 
\begin{align}
Z_t = h(V_t, t)  & = \frac{1}{\sqrt{2 \alpha \theta_0 }} \int \frac{1}{\sqrt{(v + p_t)(1 - v - p_t)}} \,dv \Bigg{\vert}_{v = V_t}   = - \sqrt{ \frac{2}{ \alpha \theta_0 }} \arcsin (\sqrt{ 1 - V_t - p_t}) \label{eq:LampZ}
\end{align}
that, after applying It\^{o}'s formula on $h(V_t, t)$, leads to the following SDE with state independent unit diffusion term
\begin{align}
dZ_t = \Bigg[  & \frac{\dot{p}_t}{ \sqrt{2 \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}}  + \frac{- \theta_t V_t}{ \sqrt{2  \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}} - \frac{1}{4} \frac{\sqrt{2 \alpha \theta_0} \left( 1 - 2 (V_t + p_t)\right)}{\sqrt{(V_t + p_t)(1 - V_t - p_t)}}  \Bigg] \,dt + dW_t \,. \label{eq:stindepSDE}
\end{align}
After replacing $V_t = 1 - p_t - \sin^2 \left(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \right) $ in (\ref{eq:stindepSDE}), we obtain that the  process $Z$ satisfies the SDE
\begin{align}
dZ_t & = \Bigg[ \frac{ \dot{p}_t  - \theta_t  \left(1 - p_t - \sin^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\sqrt{2 \alpha \theta_0} \cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)}   - \frac{1}{4}  \frac{\sqrt{2 \alpha \theta_0} \left(1 - 2 \cos^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)} \Bigg] \,dt + dW_t \nonumber \\
&  = \left[  \frac{  2  \dot{p}_t - \theta_t (1 - 2 p_t)  + (\alpha \theta_0 - \theta_t) \cos(- \sqrt{2 \alpha \theta_0 } Z_t) }{\sqrt{2 \alpha \theta_0} \sin{(- \sqrt{2 \alpha \theta_0} Z_t)}}  \right] \,dt + dW_t \,.  \label{eq:stindepSDE2}
\end{align}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/LP_t_LP.eps}
\includegraphics[width=0.45\textwidth]{plots/MP_t_LP.eps}\\
\includegraphics[width=0.45\textwidth]{plots/HP_t_LP.eps}
\includegraphics[width=0.45\textwidth]{plots/AP_t_LP.eps}
\caption{Transitions after the Lamperti transform. The transitions now present a normal Gaussian shape. This effect motivates the use of Gaussian proxies to approximate the process after using the Lamperti transform.}
  \label{fig:LP_transitions}
\end{figure}

@@

%---BEGIN SECTION 4---
\section{Likelihood in $V$ space} \label{Section_4}

\subsection{Likelihood}

@@

Closed-form solutions to initial-boundary value problem for time-in\-ho\-mo\-geneous diffusions can be obtained only in a few cases. Besides, in our case solving numerically (\ref{eq:fpk}) for the transition densities of the process $V$ at every transition step is computationally expensive. 
Therefore, under the likelihood-based inferential paradigma, many techniques have been devised to obtain approximate maximum likelihood estimates for the unknown parameters of continuous-time SDE models with discrete observations. 
 {\color{red} Rewrite: Parametric estimation problems for diffusion processes sampled at discrete times are presented in \autocite[Chapter 3]{iacus1}, and a survey of estimation methods for the parameter vector of the general one-dimensional, time-homogeneous SDE from a single sample of observations at discrete times is presented in \autocite{hurn}.}

\subsection{Approximate Likelihood}

Gaussian approximations to the transition densities of nonlinear time-in\-ho\-mo\-geneous SDEs are available through different algorithms  \autocite[Chapter 9]{saso}. However, as Figure (\ref{fig:error_transitions}) may suggest at a first glance, the choice of a Gaussian density could be inadequate when straightly applied to approximate the transition density of the forecast error $V$ of the normalized wind power production.

@@

@@

\subsection{Optimization} \label{opt_sec}

@@

To guarantee the good behave for our optimization algorithm, we aim to start the optimization as close as we can from the optimal parameters. We use \textbf{Lease Square Minimization} and \textbf{Quadratic Variation} over the data to find an initial guess, and to tune the hyper-parameter $\epsilon\in(0,\frac{1}{2})$.

@@

\item \textbf{Least Square Minimization:} We consider a set of transitions $\{\Delta V_i\}_{i}$ with $\Delta V_i=V_{i+1}-V_i$ and $\Delta t=t_{i+1}-t_i$. $(V_{i+1}|V_i)$ is a random variable which conditional mean can be approximated by the solution of the system
\begin{equation*}
\begin{cases}
\dif \E\left[V\right]&=-\theta_t\E\left[V\right]\dif t\\
\E\left[V(t_i)\right]&=V_i,
\end{cases}
\end{equation*}
evaluated in $t_{i+1}$ (i.e., $\E\left[V(t_{i+1})\right]$). Then, the random variable $(V_{i+1}-\E\left[V(t_{i+1})\right])$ has approximately zero mean. If we assume that $\theta_t=c\in\R^+$ for all $t\in[t_i,t_{i+1}]$, then $\E\left[V(t_{i+1})\right]=V_ie^{-c\Delta t}$. If we have a total of $n$ transitions, we can write the regression problem for the conditional mean with $L^2$ loss function as
\begin{equation}
c^*\approx\arg\min_{c\geq0}\left[\sum_{i=1}^n\left(V_{i+1}-\E\left[V(t_{i+1})\right]\right)^2\right]=\arg\min_{c\geq0}\left[\sum_{i=1}^n\left(V_{i+1}-V_ie^{-c\Delta t}\right)^2\right].
\label{Eq-1}
\end{equation}
As Equation (\ref{Eq-1}) is convex in $c$ (composition of convex is convex), it is enough to verify the first order optimality conditions.
\begin{equation*}
\begin{split}
\frac{\partial f}{\partial c}&=\sum_{i=1}^n2(-V_i)(-\Delta t)(V_{i+1}-V_i(1-\theta_0\Delta t))\\
&=\sum_{i=1}^n2V_i\Delta t(V_{i+1}-V_i(1-c\Delta t))\\
&=\sum_{i=1}^n2V_{i+1}V_i\Delta t-2V_i^2\Delta t+2V_i^2\Delta t^2c.
\end{split}
\end{equation*}
Then, $c^{**}$ satisfies
\begin{equation}
c^{**}\approx\frac{\sum_{i=1}^nV_i(V_i-V_{i+1})}{\Delta t\cdot\sum_{i=1}^n(V_i)^2}\quad\text{and}\quad c^*=\max\{c^{**},0\}.
\label{Eq-4}
\end{equation}
Notice that $c^{**}$ has dimension $\mathbf{time}^{-1}$.

@@

\subsubsection{Negative Log-Likelihood Minimization in $V$ Space} \label{Sec:MinLH}

@@

\subsubsection{Negative Log-Likelihood Minimization in $Z$ Space} \label{Sec:MinLHL}

@@

%---BEGIN SECTION 5---
\section{Model Comparison} \label{Section_5}

@@

\begin{table}[H]
\centering
\begin{tabular}{cccccc}
\toprule
Model & Forecast Provider & Method & Product $\theta_0\alpha$   & AIC & BIC \\ \midrule
Model 1 & Provider A & Gaussian Proxy & 0.105 & -58226 & -58211 \\
 &  & Shoji-Ozaki & 0.104 & -58226 & -58211 \\
 &  & Beta Proxy & 0.104 & -58286 & -58271 \\
 & Provider B & Gaussian Proxy & 0.105 & -58226   & -58211 \\
 &  & Shoji-Ozaki & 0.104 & -58226 & -58211 \\
 &  & Beta Proxy & 0.104 & -58288 & -58273 \\
 & Provider C & Gaussian Proxy & 0.105 & -58226 & -58211 \\
 &  & Shoji-Ozaki & 0.104 & -58226 & -58211 \\
 &  & Beta Proxy & 0.104 & -58286 & -58271 \\
Model 2 & Provider A & Beta Proxy & 0.097 & -73700   & -73685 \\ 
 & Provider B & Beta Proxy & 0.098 &  -73502 & -73487 \\ 
 & Provider C & Beta Proxy & 0.108 & -72518 & -72503 \\ 
\bottomrule
\end{tabular}
\caption{We compare the different models based on information criterion. \add }
\label{tab:model_comparison}
\end{table}
%---END SECTION 5---

%---BEGIN SECTION 6---
\section{Forecast Provider Comparison} \label{Section_6}
We compare forecasts from two different companies for the same period.
\begin{table}[H]
\centering
\begin{tabular}{ccc}
\toprule
Forecast Provider & Parameters $(\theta_0, \alpha)$ & Product $\theta_0\alpha$ \\ \midrule
Provider A  & $(1.930,0.050)$  &  0.097 \\
Provider B  & $(1.420,0.069) $  &  0.098 \\ 
Provider C  & $(1.380,0.078) $  &  0.108 \\ 
\bottomrule
\end{tabular}
\caption{Optimal parameters for different providers using Model 2 with Beta proxies.}
\label{tab:forcast_comparison}
\end{table}

%---BEGIN SECTION 7---
\section{Results} \label{Section_7}

\subsection{Initial guesses}

\subsection{Negative Log-Likelihood in $V$ Space}

\subsection{Negative Log-Likelihood in $Z$ Space}

\subsection{Main guesses}

{\color{red} Remind that Lamperti transform has the purpose tho check the consistency of the models (if the estimation of $\bm{\theta}$ with $V$ and $Z$ differs it is alarming.)}

{\color{red} Discuss to which extent it is needed more accuracy in the estimation of $\bm{\theta}$. How many data we need to achive enough accuracy for the estimates of the two parameters? Generally, such accuracy is application dependent}

 We are able to obtain the parameters based on the complete data sets. Using the different models variations, we are able to simulate wind power production given a forecast. We see in figures ()-()
five possible wind power production paths for each model.\\

In Figures ()-(), we show point-wise empirical confidence bands for the different models.

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{../../MATLAB_Files/Results/paths_testing_days/optimal_value/1.eps}
\includegraphics[width=0.45\textwidth]{../../MATLAB_Files/Results/paths_testing_days/optimal_value/2.eps}\\
\quad\\
\includegraphics[width=0.45\textwidth]{../../MATLAB_Files/Results/paths_testing_days/optimal_value/3.eps}
\includegraphics[width=0.45\textwidth]{../../MATLAB_Files/Results/paths_testing_days/optimal_value/4.eps}
\caption{We simulate following model 2  five possible future wind power production paths using the obtained optimal parameters $(\theta_0, \alpha ,\delta)=(1.2,0.1,0.6)$. Forecast is from Uruguay pertaining to  year 2019.}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{../../MATLAB_Files/Results/bands_testing_days/optimal_value/1.eps}
\includegraphics[width=0.45\textwidth]{../../MATLAB_Files/Results/bands_testing_days/optimal_value/2.eps}\\
\quad\\
\includegraphics[width=0.45\textwidth]{../../MATLAB_Files/Results/bands_testing_days/optimal_value/3.eps}
\includegraphics[width=0.45\textwidth]{../../MATLAB_Files/Results/bands_testing_days/optimal_value/4.eps}
\caption{We obtain confidence intervals following model 2  for future wind power production using the obtained optimal parameters $(\theta_0, \alpha ,\delta)=(1.2,0.1,0.6)$. Actual production plotted in retrospect. Forecast and data is from Uruguay pertaining to  year 2019.}
\end{figure}

@@

%---END SECTION 7---

%---BEGIN SECTION 8---
\section{Conclusions} \label{Section_8}

@@

We were also able to compare two different forecast providers with respect to their real-world performance on the aggregated data set and on specific wind farm sites. Finally, the model paves the way for stochastic optimal control methods enabling optimal decision making under uncertainty.
%---END SECTION 8---

>>>>

{\color{red} REWRITE In this paper we present the phenomenological model based on SDE approach in Section \ref{Section_3} and describe the physical constraints in and how these constrains can be met. Then, in Section \ref{Section_4}, we will introduce an alternative formulation of the model in Lamperti space. In Section \ref{Section_6}, we show our parameter estimation procedure and its results in Section \ref{Section_7}. We compare alternative models in Section \ref{Section_8} and different forecast providers in Section \ref{Section_9}.}

@@

%---BEGIN SECTION 2---
\section{Wind power production data in Uruguay and forecast providers } \label{Section_2}

{\color{red} Description of the global 2019 dataset, then the dataset after removing the days with curtailment (curtailment detection and removal), analysis of the dataset without curtailment.}
We have available a year long data set from Uruguay based on 363 observation paths, each of which is 24-hours long with observations recorded every 10 min. In total, it is a data set of approximately fifty thousand data points recorded in 2019. See Figure (\ref{fig:sample_data}). Data are normalized with respect to the maximum power capacity of wind power production in Uruguay, which is $\SI{1474}{\mega\watt}$. However, sometimes the real production is artificially modifies due to curtailing. This effect becomes clear when we inspect for different power production levels. We split the available data into low, medium and high power range. See Figure (\ref{fig:data_curtailing}).\\
We identify and remove the days with curtailing, removing a total of 108 days. From the resulting 255 days, we use 127 days to train the system and 128 to test it. Also, to avoid correlation between days, we intercalate the days we use for training and testing ({\color{red} Show that 24h is enough to ensure independence}).

\begin{figure}[H]
\centering
\includegraphics[width=0.35\textwidth]{plots/5.eps}
\includegraphics[width=0.35\textwidth]{plots/245.eps}\\
\includegraphics[width=0.35\textwidth]{plots/661.eps}
\includegraphics[width=0.35\textwidth]{plots/805.eps}
\caption{Four samples from the Uruguayan data of 2019. Each sample compromises of two 24-hour paths. In blue is an hourly wind power production forecast. In orange is the actual wind power production recorded in 10 minute intervals.}
  \label{fig:sample_data}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.35\textwidth]{plots/LP_6.eps}
\includegraphics[width=0.35\textwidth]{plots/MP_6.eps}\\
\includegraphics[width=0.35\textwidth]{plots/HP_6.eps}
\includegraphics[width=0.35\textwidth]{plots/AP_6.eps}
\caption{We see that forecast errors exhibit skewness, which is exaggerated as a consequence of curtailing in the production. Low power is when produced power is in $[0,0.3]$, mid-power is when it is in $(0.3,0.6]$, and high power when it is in $(0.6,1]$. We have a total of 363 days to use.}
  \label{fig:data_curtailing}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.35\textwidth]{plots/LP.eps}
\includegraphics[width=0.35\textwidth]{plots/MP.eps}\\
\includegraphics[width=0.35\textwidth]{plots/HP.eps}
\includegraphics[width=0.35\textwidth]{plots/AP.eps}
\caption{We observe that skewness has been reduced after removing the days with curtailing. We have a total of 255 days with no curtailing.}
  \label{fig:data_after_clean}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=0.35\textwidth]{plots/LP_t.eps}
\includegraphics[width=0.35\textwidth]{plots/MP_t.eps}\\
\includegraphics[width=0.35\textwidth]{plots/HP_t.eps}
\includegraphics[width=0.35\textwidth]{plots/AP_t.eps}
\caption{Transitions in error between forecast and real production. We can observe that they are not totally centered around zero.}
\label{fig:error_transitions}
\end{figure}

%---BEGIN SECTION 3---
\section{Phenomenological  Model} \label{Section_3}

@@

%---BEGIN SECTION 4---
\section{State independent diffusion term: Lamperti transform} \label{Section_4}

Our model (\ref{VtSDE}) for the forecast error has a diffusion term that depends on the state variable $V_t$. To estimate the unknown model parameters, a recommended technique is to modify the SDE (\ref{VtSDE}) by applying the so-called Lamperti transform (see \cite[40--41]{iacus1}, \cite{moma}, \cite[98--100]{saso}) to the process $V$ to obtain a SDE for the transformed process whose diffusion term does not depend anymore on the state variable. To this purpose, we consider the following Lamperti transformation  
%{\color{red} (Remind that continuous-time models with different diffusion term are singular and the likelihood cannot be written down. Besides, it can be unstable to work with low frequency. If the frequency of the data is too low, the Lamperti transform helps to detect this problem (in principle, it is not clear how to determine the right frequency for estimation. )}
%\begin{equation}
%Z_t = h(V_t, t) = \frac{1}{\sqrt{2 \theta_0 \alpha}} \int_{\xi}^{x} \frac{1}{\sqrt{(u + p_t)(1 - u - p_t)}} \,du \Bigg{\vert}_{x = V_t} \,,
%\end{equation}
%where $\xi$ is an arbitrary point of the state space of the process $V$. The choice of $\xi = \frac{1}{2} - p_t$, where $p_t$ is a known input, 
\begin{align}
Z_t = h(V_t, t; \bm{\theta} )  & = \frac{1}{\sqrt{2 \alpha \theta_0 }} \int \frac{1}{\sqrt{(v + p_t)(1 - v - p_t)}} \,dv \Bigg{\vert}_{v = V_t}   = - \sqrt{ \frac{2}{ \alpha \theta_0 }} \arcsin (\sqrt{ 1 - V_t - p_t}) \label{eq:LampZ}
\end{align}
that, after applying It\^{o}'s formula on $h(V_t, t)$, leads to the following SDE with state independent unit diffusion term
\begin{align}
dZ_t = \Bigg[  & \frac{\dot{p}_t}{ \sqrt{2 \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}}  + \frac{- \theta_t V_t}{ \sqrt{2  \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}} - \frac{1}{4} \frac{\sqrt{2 \alpha \theta_0} \left( 1 - 2 (V_t + p_t)\right)}{\sqrt{(V_t + p_t)(1 - V_t - p_t)}}  \Bigg] \,dt + dW_t \,. \label{eq:stindepSDE}
\end{align}
After replacing $V_t = 1 - p_t - \sin^2 \left(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \right) $ in (\ref{eq:stindepSDE}), we obtain that the  process $Z$ satisfies the SDE
\begin{align}
dZ_t & = \Bigg[ \frac{ \dot{p}_t  - \theta_t  \left(1 - p_t - \sin^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\sqrt{2 \alpha \theta_0} \cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)}   - \frac{1}{4}  \frac{\sqrt{2 \alpha \theta_0} \left(1 - 2 \cos^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)} \Bigg] \,dt + dW_t \nonumber \\
&  = \left[  \frac{  2  \dot{p}_t - \theta_t (1 - 2 p_t)  + (\alpha \theta_0 - \theta_t) \cos(- \sqrt{2 \alpha \theta_0 } Z_t) }{\sqrt{2 \alpha \theta_0} \sin{(- \sqrt{2 \alpha \theta_0} Z_t)}}  \right] \,dt + dW_t \,.  \label{eq:stindepSDE2}
\end{align}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.35\textwidth]{plots/LP_t_LP.eps}
\includegraphics[width=0.35\textwidth]{plots/MP_t_LP.eps}\\
\includegraphics[width=0.35\textwidth]{plots/HP_t_LP.eps}
\includegraphics[width=0.35\textwidth]{plots/AP_t_LP.eps}
\caption{Transitions after the Lamperti transform. The transitions now present a normal Gaussian shape. This effect motivates the use of Gaussian proxies to approximate the process after using the Lamperti transform.}
  \label{fig:LP_transitions}
\end{figure}

@@

%{\color{red} Remind that Lamperti transform has the purpose tho check the consistency of the models, that is if the estimation of $\bm{\theta}$ with $V$ and $Z$ differs it is alarming.}

%---BEGIN SECTION 5---
\section{ Likelihood functions of the forecast errors data and optimization algorithm} \label{Section_5}

\subsection{Likelihood in the $V-$space}

@@

Closed-form solutions to initial-boundary value problem for time-in\-ho\-mo\-geneous diffusions can be obtained only in a few cases (see, for example, \autocite[Section 3.1]{eglix}). Besides, in our case solving numerically (\ref{eq:fpk}) for the transition densities of the process $V$ at every transition step is computationally expensive. 
Therefore, under the likelihood-based inferential paradigma, many techniques have been devised to obtain approximate maximum likelihood estimates for the unknown parameters of continuous-time SDE models with discrete observations. 
 {\color{red} Rewrite: Parametric estimation problems for diffusion processes sampled at discrete times are presented in \autocite[Chapter 3]{iacus1}, consider Preston and Wood, then Egorov, and then Sorensen, remove the
 survey of estimation methods for the parameter vector of the general one-dimensional, time-homogeneous SDE from a single sample of observations at discrete times is presented in \autocite{hurn}, refer to time-homogeneous.}

\subsection{Approximate likelihood  in the $V-$space}

Gaussian approximations to the transition densities of nonlinear time-in\-ho\-mo\-geneous SDEs are available through different algorithms  \autocite[Chapter 9]{saso}. However, as Figure \ref{fig:error_transitions} may suggest at a first glance, the choice of a Gaussian density could be inadequate when straightly applied to approximate the transition density of the forecast error $V$ of the normalized wind power production.

@@

\subsection{Approximate likelihood  in the $Z-$space}

@@

\subsection{Algorithm for the approximate maximum likelihood estimations} \label{opt_sec}

@@

To guarantee the good behave for our optimization algorithm, we aim to start the optimization as close as we can from the optimal parameters. We use \textbf{Least Square Minimization} and \textbf{Quadratic Variation} over the data to find an initial guess, and to tune the hyper-parameter $\epsilon\in(0,\frac{1}{2})$.

@@

\item \textbf{Least Square Minimization:} We consider a set of transitions $\{\Delta V_i\}_{i}$ with $\Delta V_i=V_{i+1}-V_i$ and $\Delta t=t_{i+1}-t_i$. $(V_{i+1}|V_i)$ is a random variable which conditional mean can be approximated by the solution of the system
\begin{equation*}
\begin{cases}
\dif \E\left[V\right]&=-\theta_t\E\left[V\right]\dif t\\
\E\left[V(t_i)\right]&=V_i,
\end{cases}
\end{equation*}
evaluated in $t_{i+1}$ (i.e., $\E\left[V(t_{i+1})\right]$). Then, the random variable $(V_{i+1}-\E\left[V(t_{i+1})\right])$ has approximately zero mean. If we assume that $\theta_t=c\in\R^+$ for all $t\in[t_i,t_{i+1}]$, then $\E\left[V(t_{i+1})\right]=V_ie^{-c\Delta t}$. If we have a total of $n$ transitions, we can write the regression problem for the conditional mean with $L^2$ loss function as
\begin{equation}
c^*\approx\arg\min_{c\geq0}\left[\sum_{i=1}^n\left(V_{i+1}-\E\left[V(t_{i+1})\right]\right)^2\right]=\arg\min_{c\geq0}\left[\sum_{i=1}^n\left(V_{i+1}-V_ie^{-c\Delta t}\right)^2\right].
\label{Eq-1}
\end{equation}
As Equation (\ref{Eq-1}) is convex in $c$ (composition of convex is convex), it is enough to verify the first order optimality conditions. From
\begin{equation*}
\begin{split}
\frac{\partial f}{\partial c}&=\sum_{i=1}^n2(-V_i)(-\Delta t)(V_{i+1}-V_i(1-\theta_0\Delta t))
=\sum_{i=1}^n2V_i\Delta t(V_{i+1}-V_i(1-c\Delta t))\\
&=\sum_{i=1}^n2V_{i+1}V_i\Delta t-2V_i^2\Delta t+2V_i^2\Delta t^2c.
\end{split}
\end{equation*}
it follows that $c^{**}$ satisfies
\begin{equation}
c^{**}\approx\frac{\sum_{i=1}^nV_i(V_i-V_{i+1})}{\Delta t\cdot\sum_{i=1}^n(V_i)^2}\quad\text{and}\quad c^*=\max\{c^{**},0\}.
\label{Eq-4}
\end{equation}
Notice that $c^{**}$ has dimension $\mathbf{time}^{-1}$.

@@

\subsubsection{Negative log-likelihood minimization in the $V-$space} \label{Sec:MinLH}

@@

\subsubsection{Negative log-likelihood minimization in the $Z-$space} \label{Sec:MinLHL}

@@

%---END SECTION 5---

%---BEGIN SECTION 6---
\section{Applications to the Uruguay wind and forecast dataset} \label{Section_6}

\subsection{Calibration of the approximate negative log-likelihood in the $V-$space}

\subsection{Calibration of the approximate negative log-likelihood in the $Z-$space}

\subsection{Model comparison} \label{Model_Comp}

@@

\begin{table}[H]
\centering
\begin{tabular}{cccccc}
\toprule
Model & Forecast Provider & Method & Product $\theta_0\alpha$   & AIC & BIC \\ \midrule
Model 1 & Provider A & Gaussian Proxy & 0.105 & -58226 & -58211 \\
 &  & Shoji-Ozaki & 0.104 & -58226 & -58211 \\
 &  & Beta Proxy & 0.104 & -58286 & -58271 \\
 & Provider B & Gaussian Proxy & 0.105 & -58226   & -58211 \\
 &  & Shoji-Ozaki & 0.104 & -58226 & -58211 \\
 &  & Beta Proxy & 0.104 & -58288 & -58273 \\
 & Provider C & Gaussian Proxy & 0.105 & -58226 & -58211 \\
 &  & Shoji-Ozaki & 0.104 & -58226 & -58211 \\
 &  & Beta Proxy & 0.104 & -58286 & -58271 \\
Model 2 & Provider A & Beta Proxy & 0.097 & -73700   & -73685 \\ 
 & Provider B & Beta Proxy & 0.098 &  -73502 & -73487 \\ 
 & Provider C & Beta Proxy & 0.108 & -72518 & -72503 \\ 
\bottomrule
\end{tabular}
\caption{We compare the different models based on information criterion. \add }
\label{tab:model_comparison}
\end{table}

\subsection{Forecast providers comparison} \label{Forecast_Comp}
We compare forecasts from two different companies for the same period.
\begin{table}[H]
\centering
\begin{tabular}{ccc}
\toprule
Forecast Provider & Parameters $(\theta_0, \alpha)$ & Product $\theta_0\alpha$ \\ \midrule
Provider A  & $(1.930,0.050)$  &  0.097 \\
Provider B  & $(1.420,0.069) $  &  0.098 \\ 
Provider C  & $(1.380,0.078) $  &  0.108 \\ 
\bottomrule
\end{tabular}
\caption{Optimal parameters for different providers using Model 2 with Beta proxies.}
\label{tab:forcast_comparison}
\end{table}

{\color{red} Discuss to which extent it is needed more accuracy in the estimation of $\bm{\theta}$. How much data do we need to achieve enough accuracy for the estimates of the two parameters? Generally, such accuracy is application dependent.}

\subsection{Calibration of Model 2 with additional parameter $\delta$ TO DO}

After calibrating Model 2 on the training set, we can generate simulations of the wind power production for the time horizon of interest. The next Figure \ref{fig:simulation_paths} shows five simulated paths of the wind power production for each day of interest, assuming Model 2 specification and a given forecaster.

\begin{figure}[H]
\centering
\includegraphics[width=0.18\textwidth]{../../MATLAB_Files/Results/paths_testing_days/optimal_value/1.eps}
\includegraphics[width=0.18\textwidth]{../../MATLAB_Files/Results/paths_testing_days/optimal_value/2.eps}\\
\quad\\
\includegraphics[width=0.18\textwidth]{../../MATLAB_Files/Results/paths_testing_days/optimal_value/3.eps}
\includegraphics[width=0.18\textwidth]{../../MATLAB_Files/Results/paths_testing_days/optimal_value/4.eps}
\caption{Five simulated wind power production paths using the approximate MLEs for Model 2 $(\theta_0, \alpha ,\delta)=(1.2,0.1,0.6)$. }
\label{fig:simulation_paths}
\end{figure}

On the basis of the approximate MLEs for Model 2, we can obtain empirical pointwise confidence bands for the wind power production. The next Figure \ref{fig:confidence_bands} shows the empirical pointwise confidence bands for the wind power production for each day of interest, assuming Model 2 specification and a given forecaster.

\begin{figure}[H]
\centering
\includegraphics[width=0.35\textwidth]{../../MATLAB_Files/Results/bands_testing_days/optimal_value/1.eps}
\includegraphics[width=0.35\textwidth]{../../MATLAB_Files/Results/bands_testing_days/optimal_value/2.eps}\\
\quad\\
\includegraphics[width=0.35\textwidth]{../../MATLAB_Files/Results/bands_testing_days/optimal_value/3.eps}
\includegraphics[width=0.35\textwidth]{../../MATLAB_Files/Results/bands_testing_days/optimal_value/4.eps}
\caption{Empirical pointwise confidence bands for the wind power production using the approximate MLEs for Model 2 $(\theta_0, \alpha ,\delta)=(1.2,0.1,0.6)$. Blue line: real production.}
\label{fig:confidence_bands}
\end{figure}

@@

%---BEGIN SECTION 7---
\section{Conclusions} \label{Section_7}

@@

We were also able to compare two different forecast providers with respect to their real-world performance on the aggregated data set and on specific wind farm sites. Finally, the model paves the way for stochastic optimal control methods enabling optimal decision making under uncertainty.
%---END SECTION 7---
